---
title: "SOC 506: Analysis and cleaning "
output: html_document

---

```{r, include = F}
# Load packages 
library(readr)
library(tidyverse)
library(effects)
library(gtsummary)
```

```{r}
data_gss <- read_csv("../Data/gss_proj_data.csv")

# Data cleaning. 
str(data_gss)

gss_clean <- data_gss %>% 
  filter(sexeduc != "NA")

  gss_clean <- gss_clean %>%  filter(income %in% c(1:13),
                        degree %in% (0:4)) %>%
    mutate(sexeduc = plyr :: mapvalues(sexeduc, from = c(1,2,3), to = c("favor","oppose","depends")),
           race = plyr :: mapvalues(race, from = c(1,2,3), to = c("white","black","other")), 
           sex = plyr :: mapvalues(sex, from = c(1,2), to = c("male","female")),
           income = plyr :: mapvalues(income, 
                                      from = c(1:13),
                                      to = c("1,000 or less",
                                             "1,000 to 2,999",
                                             "3,000 to 3,999",
                                             "4,000 to 4,999",
                                             "5,000 to 5,999",
                                             "6,000 to 6,999",
                                             "7,000 to 7,999",
                                             "8,000 to 9,999",
                                             "10,000 to 14,999",
                                             "15,000 to 19,999",
                                             "20,000 to 24,999",
                                             "25,000 or more",
                                             "Refused")),
           rincome = plyr :: mapvalues(rincome, from = c(1:13), 
                                       to = c("1,000 or less",
                                             "1,000 to 2,999",
                                             "3,000 to 3,999",
                                             "4,000 to 4,999",
                                             "5,000 to 5,999",
                                             "6,000 to 6,999",
                                             "7,000 to 7,999",
                                             "8,000 to 9,999",
                                             "10,000 to 14,999",
                                             "15,000 to 19,999",
                                             "20,000 to 24,999",
                                             "25,000 or more",
                                             "Refused")),
           degree = plyr :: mapvalues(degree, 
                                      from= c(0:4),
                                      to = c("Less than HS",
                                             "HS",
                                             "JR college",
                                             "Bachelor",
                                             "Graduate")),
           hispanic = ifelse(hispanic == 1,"Not Hispanic","Hispanic"),
           region = plyr :: mapvalues(region, from = c(1:9),
                                      to = c("New England", 
                                             "Mid Atlantic",
                                             "E N Central",
                                             "W N Central",
                                             "S Atlantic",
                                             "E S Central",
                                             "W S Central",
                                             "Mountain",
                                             "Pacific"))) %>% 
    select(year,race,sex,sexeduc,hispanic,region,income,rincome,degree,cohort) 
  # Need to fix from haven to factor variables. 
## Clean GSS data
gss_clean <- gss_clean %>% filter(rincome != is.na(rincome))

```


```{r}
# Look at the data. 
# Create some summary statistics 
gss_clean <- gss_clean %>% na.omit()
# Write GSS clean csv 
#write_csv(gss_clean, file = "../Data/gss_clean.csv")

data <- read_csv("../Data/gss_clean.csv")
```

```{r}
# logistic regression model 
data_logistic <- data %>% mutate(sex_ed_approval = ifelse(sexeduc == "favor",1,0)) %>% 
  na.omit

# GLM Base. 
glm_base <- glm(sex_ed_approval ~ rincome, 
                data = data_logistic, 
                family = binomial(link = "logit"))
summary(glm_base) # Reference category is the lowest income level. 
```

For those that are in the 25,000 dollar a year or more category, thre is a significant `r summary(glm_base)$coefficients[6]` increase in the log odds of approving of sex education in public schools. All of the other categories are not significant. 

```{r}
# Full glm 
glm_full <- glm(sex_ed_approval ~ rincome + race + sex + 
                  degree + hispanic + cohort, 
                  data = data_logistic,
    family = binomial(link = "logit"))
summary(glm_full)
```

Adding in the other predictors, controlling for race, gender, and year born, all are insignificant except for cohort. For a one year increase in the year born, there is a `r summary(glm_full)$coefficients[29]` increase in the log-odds of favoring sex education in public schools holding all else constant. 

```{r}
# Create nice coefficient names. 

coefs = c("(Intercept)" = "Constant",
             "rincome1,000 to 2,999" = "$1000-2,999",
             "rincome3,000 to 3,999" = "$3000-3,999",
             "rincome4,000 to 4,999" = "$4,000-4,999",
             "rincome5,000 to 5,999" = "$5,000-5,999",
             "rincome6,000 to 6,999" = "$6,000-6,999",
             "rincome7,000 to 7,999" = "$7,000-7,999",
             "rincome8,000 to 9,999" = "8,000-9,999",
             "rincome10,000 to 14,999" = "$10,000-14,999",
             "rincome15,000 to 19,999" = "$15,000-19,999",
             "rincome20,000 to 24,999" = "$20,000-24,999",
             "rincome25,000 or more" = "$25,000 or more",
             "raceother" = "Race: Other",
             "racewhite" = "Race:White",
             "sexmale" = "Male",
             "degreeGraduate" = "Graduate Degree",
             "degreeHS" = "High School",
             "degreeJR college" = "Community College",
             "degreeLess than HS" = "less than HS",
             "hispanicNot Hispanic" = "Not Hispanic",
             "cohort" = "Year born")
modelsummary:: modelsummary(list(glm_base,glm_full),
                            coef_map = coefs, 
                            estimate = "{estimate}{stars}", 
                            title = "Full and Restricted Logistic Models (Log Odds)")

```

## Model Checking 

## LRT 

```{r}
## LRT 
lmtest::lrtest(glm_base,glm_full)
```

Running a model comparison, it is clear from a clear statistically significant difference in the log-likelihoods. In addition the AIC and BIC are lower for the full model. 

##  Confusion Matrix 

```{r}
library(caret)

# Rep data 
mydata2<- data_logistic
# convert admit to a factor 
mydata2$sex_ed_approval<-factor(mydata2$sex_ed_approval,
                      # levels are 0,1 to disapprove and approve.
                      levels=c("0","1"),labels=c("Disapprove","Approve"))
# Partition the data into a trianing and test set 
Train <- createDataPartition(mydata2$sex_ed_approval, 
                             # calssify and admit as .6
                             p=0.6, list=FALSE)
# create the training set 
training <-data.frame(mydata2[ Train[,1], ])
# create the test set
testing <- data.frame(mydata2[ -Train[,1], ])
# Fit the logit model 
mod_fit <- train(sex_ed_approval ~ rincome + race + sex + 
                  degree + hispanic + cohort,
                 data=training, method="glm", family="binomial")
# predict with the new data 
pred <- predict(mod_fit, newdata=testing)
# compute the accuracy score 
accuracy <- table(pred, testing[,"sex_ed_approval"])
# out put the sim of the diagonals of accuracy over the sum of accuracy 
cat('Accuracy:',sum(diag(accuracy))/sum(accuracy),'\n')
# predict  the model again
pred <- predict(mod_fit, newdata=testing)
# create a confusion matrix 
confusionMatrix(data=pred, testing$sex_ed_approval)
```

The confusion matrix shows that the model did not classify any of the people as disapproving, and all of the as approving, this is potentially a problem since we know that there is variation. In total, there was a negative predicted value of .93 which indicates that the model is really good at lassifyong people as approving but pretty bad otherwise.  

# Visuals 

```{r}
# Create a visual of the predicted probabilities 
newdata <- with(data_logistic, data.frame(
  # Set all coefficients to constant, except cohort and sex. 
  rincome = rep("25,000 or more",138),
  sex = as.factor(rbind(rep("male",
                            69),rep("female",69))),
  cohort = c(seq(1932,2000,by = 1),seq(1932,2000,by = 1)),
  race = rep("white",138),
  degree = rep("Graduate",138),
  hispanic = rep("Not Hispanic",138)
))

## genarate the predicted probabilities 
predict <- cbind(newdata,predict(glm_full,newdata,type = "link",se = T))

## genarating the predicted probability and the upper and lower bounds. 
predict <- within(predict, {predictedprobs <- plogis(fit)
LL <- plogis(fit - (1.96 * se.fit))
UL <- plogis(fit + (1.96 * se.fit))
})
predict
## now we can plot the data. 

predict %>% ggplot(aes(x = cohort,y = predictedprobs, fill = sex))  + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = sex), alpha = .2) + 
  geom_line(aes(col = sex), size = 1) + theme_minimal() + 
  labs(x = "Year Born",
     y = "Probability of Support",
    title = "Probability of supporting Sex Ed in Schools",
      subtitle = "For $ > 25,000, non-hispanic, White") + 
  scale_fill_manual(name = "Sex",
                    labels = c("Female","Male"), 
                    values = c("Orange","blue")) + 
  scale_color_manual(name = "Sex",
                    labels = c("Female","Male"), 
                    values = c("Orange","blue"))


```



